Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-14 csf-2017/paper.tex
Moved program example g1 and f3 to the same figure.

@@

\subsection{Well-bracketed control-flow} 
Using the stack-based calling convention of \texttt{\footnotesize{scall}}, we get
well-bracketed control-flow. To illustrate this, we look at
two example programs \texttt{\footnotesize{f3}} in \figurename~\ref{fig:prog-f3} and
\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-g1}.

In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order for the
assertion in the middle to succeed, they need to be well-bracketed. If the
adversary was able to store the return pointer from the first call and invoke
it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on top of its stack and
the assertion will fail. However, the security measures of our calling
convention explained in Section~\ref{sec:stack-and-return-pointer} prevent this
sort of attack: specifically, the return pointer is local, so it can only be
stored on the stack, but the part of the stack that is accessible to the
adversary is cleared before the second invocation. The following lemma shows
that there are also no other attacks that will break well-bracketedness of this
example, i.e.\ the assertion never fails. It is similar to the two previous
lemmas:
\begin{figure}[t]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

@@

The final example, in \figurename~\ref{fig:prog-g1}, is a faithful translation
of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}. \texttt{\footnotesize{g1}} is a closure generator that
generates closures with one variable $x$ set to $0$ in its environment and
\texttt{\footnotesize{f4}} as the program. \texttt{\footnotesize{f4}} expects one argument, a callback. The
program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}
sets $x$ to $1$ and calls the callback a second time. When it returns from the
callback, it asserts that $x$ is $1$ and returns. This example is more
complicated than the previous ones because it involves a closure invoked by the
adversary and an adversary callback invoked by us. As explained in
Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
to properly check the stack pointer received by our closure from the adversary
(it must have the correct permission and fit with our stack discipline) and (2)
require that the adversary callback is global.

\begin{figure}[t]
  \centering
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \caption{A program that relies on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-g1}
\end{figure}

@@

%\bibliographystyle{IEEEtran}
\bibliographystyle{plainnat}
\bibliography{IEEEabrv,references}
\end{document}
\grid

>>>>

% Include the following to reduce room around display equations
%\setlength{\abovedisplayskip}{3pt}
%\setlength{\belowdisplayskip}{3pt}
%\setlength{\abovedisplayshortskip}{3pt}
%\setlength{\belowdisplayshortskip}{3pt}

@@

\subsection{Well-bracketed control-flow} 
Using the stack-based calling convention of \texttt{\footnotesize{scall}}, we get
well-bracketed control-flow. To illustrate this, we look at
two example programs \texttt{\footnotesize{f3}} and
\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-f3-and-g1}.

In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order for the
assertion in the middle to succeed, they need to be well-bracketed. If the
adversary was able to store the return pointer from the first call and invoke
it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on top of its stack and
the assertion will fail. However, the security measures of our calling
convention explained in Section~\ref{sec:stack-and-return-pointer} prevent this
sort of attack: specifically, the return pointer is local, so it can only be
stored on the stack, but the part of the stack that is accessible to the
adversary is cleared before the second invocation. The following lemma shows
that there are also no other attacks that will break well-bracketedness of this
example, i.e.\ the assertion never fails. It is similar to the two previous
lemmas:

\begin{figure}[t]
  \centering
  \begin{minipage}[t]{5.05cm}
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{3.4cm}
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \end{minipage}
  \caption{ Two programs that rely on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-f3-and-g1}
\end{figure}

@@

The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}}, is a faithful translation
of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}. \texttt{\footnotesize{g1}} is a closure generator that
generates closures with one variable $x$ set to $0$ in its environment and
\texttt{\footnotesize{f4}} as the program. \texttt{\footnotesize{f4}} expects one argument, a callback. The
program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}
sets $x$ to $1$ and calls the callback a second time. When it returns from the
callback, it asserts that $x$ is $1$ and returns. This example is more
complicated than the previous ones because it involves a closure invoked by the
adversary and an adversary callback invoked by us. As explained in
Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
to properly check the stack pointer received by our closure from the adversary
(it must have the correct permission and fit with our stack discipline) and (2)
require that the adversary callback is global.

@@

%\bibliographystyle{IEEEtran}
\bibliographystyle{plainnat}
\bibliography{IEEEabrv,references}

\end{document}
\grid
